---
title: "Stats!"
author: "Gregory J. Matthews"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Probability 
  - What is it?  
  - Some rules
  
$P(A^c) = 1 - P(A)$

$P(A \cup B) = P(A) + P(B) - P(A \cap B)$

$P(A \cap B) = P(A) P(B)$ if $A \perp B$

$P(A) = P(A|B)P(B) + P(A|B^c)P(B)$ 

$P(A | B) = \frac{P(A\cap B)}{P(A)}$ 

$P(A\cap B) = P(A|B)P(B)$

$P(B | A) = \frac{P(A|B)P(B)}{P(A|B)P(B) + P(A|B^c)P(B^c)}$ 

## Distributions
  - pdf vs cdf
  - Normal
  - Binomial
  - Others

CDF of X: $F(x) = P(X \le x)$

PDF of X: $f(x) = P(X = x)$ ONLY IN DISCRETE CASE!

```{r}
x <- seq(-4,4,0.01)
plot(x, dnorm(x), type = "l", main = "pdf Normal")
plot(x, pnorm(x), type = "l", main = "cdf Normal")
```
```{r}
x <- c(0:10)
plot(x, dbinom(x,10,0.5),  main = "pdf binomial")
x <- seq(0,10,0.01)
plot(x, pbinom(x,10,0.5), type = "l",  main = "cdf binomial")
```
```{r}
# Let X ~ normal(0,1)
# Let Y ~ binomial(10,0.5)
# Let B ~ binomial(1,0.5)
# Let Z | B ~ normal(B, 1)
#Find probability that X > 1 give X > 0.  
#That is find P(X > 1 | X > 0). 

#Find P(Y = 5 AND -1 < X < 1)

# Find P(Z > 0)

```


## Inference
  - Point Estimates
  - Confidence intervals
  - Hypothesis testing
    - Permutation testing
    - t-test
    - $\chi^2$-squared test
    - p-values!
    
```{r}
#Lets do some simulations about point estimates and look at their properties 
#sample mean vs sample median

#Let's look at confidence interval coverage for a proportion. 

# Let's do hypothesis testing "by hand".  We'll do a t-test by hand.

#Let's do a permutation test

# Let's talk about p-values.  



```
   


