---
title: "Data Wrangling on one table"
author: "Gregory J. Matthews"
date: "8/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidyverse grammar
- select(): Returns a sub set of columns (i.e. features or variables)
- filter(): Returns a subset of rows (i.e. observations)
- mutate(): add or modify existing columns
- arrange(): sort the rows
- summarize(): aggregate the data across rows (e.g., group it according to some criteria)

#Filter and Select
```{r}
library(tidyverse)
library(mdsr)
#Subset of rows
filter(CIACountries, pop > 1000000000)

#I can also do it like this
CIACountries %>% filter(pop > 1e9)

#In base R
CIACountries[CIACountries$pop > 1e9,]

#or
subset(CIACountries, pop > 1e9)

```


```{r}
#Subset of columns
select(CIACountries, country, pop, gdp)

#I can also do it like this
CIACountries %>% select(country, pop, gdp)

#In base R
CIACountries[, c("country", "pop", "gdp")]

#or
subset(CIACountries, select = c("country", "pop", "gdp"))

```


```{r}
#filter and select
filter(select(CIACountries, country, pop, gdp), pop > 1e9)

#Or we can do this with a pipeline
CIACountries %>% select(country, pop, gdp) %>% filter(pop > 1e9)

#Or in base R
CIACountries[CIACountries$pop > 1e9,c("country", "pop", "gdp")]

#or this:
subset(CIACountries, pop > 1e9, select = c("country", "pop", "gdp"))
```
#Mutate and Rename
```{r}
#Mutate
#Population Density
my_CIACountries <- mutate(CIACountries, dens = pop/area)
my_CIACountries[1,]

my_CIACountries <- CIACountries %>% mutate(dens = pop/area)
my_CIACountries[1,]

#Base R
library(mdsr)
my_CIACountries$dens <- CIACountries$pop / CIACountries$area

#rename
my_CIACountries <- CIACountries %>% rename(education = educ)
my_CIACountries[1,]
```

#arrange

```{r}
#Sorting with arrange
#Defaults to ascending
CIACountries_sort <- arrange(CIACountries, gdp)
CIACountries_sort[1:5,]

#Sort descending
CIACountries_sort <- arrange(CIACountries, desc(gdp))
CIACountries_sort[1:5,]

CIACountries_sort <- CIACountries %>% arrange(desc(gdp))
CIACountries_sort[1:5,]

#Base R with order
CIACountries_sort <- CIACountries[order(CIACountries$gdp),]
CIACountries_sort[1:5,]

CIACountries_sort <- CIACountries[order(-CIACountries$gdp),]
CIACountries_sort[1:5,]
```

#Summarize and Group_by
```{r}
#What is the average area by high vs low education countries?
my_CIACountries <-
  CIACountries %>% mutate(high_educ = educ > 4.5) %>% group_by(high_educ) %>% summarize(mn_area = mean(area))
my_CIACountries

#Base R
my_CIACountries <- CIACountries
my_CIACountries$high_educ <- CIACountries$educ > 4.5
tapply(my_CIACountries$area,my_CIACountries$high_educ, mean)

```

##In class assignment: What first letter has the largest population? 

#What proportion of countries produce any oil stratified by net users class? 




  





