---
title: "Data Visualization"
author: "Gregory J. Matthews"
date: "8/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mdsr)
library(tidyverse)
data("CIACountries")
```

# Scatterplots
### Base R

```{r}
plot(CIACountries$gdp, 
     CIACountries$educ)
```

Add axis labels
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education")
```

Change the shape
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16)
```

Change the color
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red")
```

Change the x and y limits 
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red", 
     xlim = c(0,150000), 
     ylim = c(0,15))
```

Change the x and y limits 
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red", 
     xlim = c(0,150000), 
     ylim = c(0,15))
```

Aspect Ratio
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red", 
     xlim = c(0,150000), 
     ylim = c(0,15), 
     asp = 10000)
```

Change axis labels 
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red", 
     xlim = c(0,150000), 
     ylim = c(0,15), 
     asp = 10000, 
     xaxt = 'n', 
     yaxt = 'n')

axis(1, c(0,100000,200000), c("None","A little","A lot"))
axis(2, c(0,5,10,15), c("None","Ok","Good","Great"))
```

Size  
```{r}
plot(CIACountries$gdp, 
     CIACountries$educ, 
     xlab = "GDP",
     ylab = "Education",
     pch = 16, 
     col = "red", 
     xlim = c(0,150000), 
     ylim = c(0,15), 
     asp = 10000, 
     xaxt = 'n', 
     yaxt = 'n', 
     cex = 0.5)

axis(1, c(0,100000,200000), c("None","A little","A lot"))
axis(2, c(0,5,10,15), c("None","Ok","Good","Great"))
```

### ggplot
The key to using ggplot to is to think of ever command as a layer!

```{r}
library(ggplot2)
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point()
```

Add axis labels
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point() + 
  xlab("GDP") + 
  ylab("Education")
```

Change the shape
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23) +
  xlab("GDP") +
  ylab("Education")
```

Change the color
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23, colour = "red") +
  xlab("GDP") +
  ylab("Education") 
```

Change the x and y limits 
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23, colour = "red") +
  xlab("GDP") +
  ylab("Education") + 
  xlim(0,200000) + 
  ylim(0,15)
```

Aspect Ratio
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23, colour = "red") +
  xlab("GDP") +
  ylab("Education") + 
  xlim(0,200000) + 
  ylim(0,15) + 
  coord_fixed(ratio=10000)
```

Change axis values
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23, colour = "red") +
  xlab("GDP") +
  ylab("Education") + 
  xlim(0,200000) + 
  ylim(0,15) + 
  coord_fixed(ratio=10000) + 
  scale_x_continuous(breaks = c(0,100000,200000),
                     labels = c("None","A little","A lot"), 
                     limits = c(0,200000)) + 
  scale_y_continuous(breaks = c(0,5,10,15),
                     labels = c("None","Ok","Good","Great"),
                     limits = c(0,15)
                     )
#Note: scale_x_continuous overrules the xlim!
```

Size  
```{r}
ggplot(aes(x = gdp, y = educ), data = CIACountries) +
  geom_point(shape = 23, colour = "red", size = 0.5) +
  xlab("GDP") +
  ylab("Education") + 
  xlim(0,200000) + 
  ylim(0,15) + 
  coord_fixed(ratio=10000) + 
  scale_x_continuous(breaks = c(0,100000,200000),
                     labels = c("None","A little","A lot"), 
                     limits = c(0,200000)) + 
  scale_y_continuous(breaks = c(0,5,10,15),
                     labels = c("None","Ok","Good","Great"),
                     limits = c(0,15)
                     )

```

# Univariate displays

### Histograms

```{r}
library(mdsr)
data("SAT_2010")
```
```{r}
#Base R
hist(SAT_2010$math)
#ggplot
ggplot(aes(x = math), data = SAT_2010) + geom_histogram()
```


```{r}
#Base R
dens <- density(SAT_2010$math)
plot(dens$x, dens$y, type = "l")
#ggplot
#This is a lot cleaner code and nicer looking!
ggplot(aes(x = math), data = SAT_2010) + geom_density()
```

```{r}
sub <- head(SAT_2010, 10)
sub <- sub[order(sub$math),]
barplot(sub$math, names.arg = sub$state)

#ggplot
ggplot(
  data = head(SAT_2010, 10), 
  aes(x = reorder(state, math), y = math)
) +
  geom_col() +
  labs(x = "State", y = "Average math SAT score")
```
### Stacked bar chart
```{r}
library(mosaicData)
ggplot(data = mosaicData::HELPrct, aes(x = homeless)) + 
  geom_bar(aes(fill = substance), position = "fill") 

ggplot(data = mosaicData::HELPrct, aes(x = homeless)) + 
  geom_bar(aes(fill = substance), position = "fill") +
  scale_fill_brewer(palette = "Spectral") 


ggplot(data = mosaicData::HELPrct, aes(x = homeless)) + 
  geom_bar(aes(fill = substance), position = "fill") +
  scale_fill_brewer(palette = "Spectral") + 
  coord_flip()

```


### Multivariate displays
```{r}

g <- ggplot(
  data = SAT_2010, 
  aes(x = expenditure, y = math)
) + 
  geom_point()

```
 We can easily add a trend line with ggplot.  
```{r}
g <- g + 
  geom_smooth(method = "loess", se = FALSE) + 
  xlab("Average expenditure per student ($1000)") +
  ylab("Average score on math SAT")
```

To do this in base R you have to do this: 
```{r}
plot(SAT_2010$expenditure, SAT_2010$math)
a <- loess(SAT_2010$math ~ SAT_2010$expenditure)
points(a$x[order(a$x)], a$fitted[order(a$x)], type = "l", col = "red")
```

```{r}
SAT_2010 <- SAT_2010 %>%
  mutate(
    SAT_rate = cut(
      sat_pct, 
      breaks = c(0, 30, 60, 100), 
      labels = c("low", "medium", "high")
    )
  )
g <- g %+% SAT_2010
g + aes(color = SAT_rate)
#To this in base R is possible, but it is a pain!
```

### Faceting

```{r}
#Base R
par(mfrow = c(1, 3))

plot(SAT_2010$expenditure[SAT_2010$SAT_rate == "low"],
     SAT_2010$math[SAT_2010$SAT_rate == "low"],
     xlab = "expenditure",
     ylab = "math score")

plot(SAT_2010$expenditure[SAT_2010$SAT_rate == "medium"],
     SAT_2010$math[SAT_2010$SAT_rate == "medium"],
     xlab = "expenditure",
     ylab = "math score")

plot(SAT_2010$expenditure[SAT_2010$SAT_rate == "high"],
     SAT_2010$math[SAT_2010$SAT_rate == "high"],
     xlab = "expenditure",
     ylab = "math score")
```

So much easier in ggplot!

```{r}
g + facet_wrap(~ SAT_rate)
```

```{r}
#full code
SAT_2010 <- SAT_2010 %>%
  mutate(SAT_rate = cut(
    sat_pct,
    breaks = c(0, 30, 60, 100),
    labels = c("low", "medium", "high")
  ))
g <- ggplot(data = SAT_2010,
            aes(x = expenditure, y = math)) +
  geom_point() + geom_smooth(method = "lm") + facet_wrap( ~ SAT_rate)
g

```

NHANES example
```{r}
library(NHANES)
ggplot(
  data = slice_sample(NHANES, n = 1000), 
  aes(x = Age, y = Height, color = Gender)
) + 
  geom_point() + 
  geom_smooth() + 
  xlab("Age (years)") + 
  ylab("Height (cm)") +
  labs(color = "Gender")
```

```{r}
#relevel the reference category
library(NHANES)
ggplot(
  data = slice_sample(NHANES, n = 1000), 
  aes(x = Age, y = Height, color = fct_relevel(Gender, "male"))
) + 
  geom_point() + 
  geom_smooth() + 
  xlab("Age (years)") + 
  ylab("Height (cm)") +
  labs(color = "Gender")
```

```{r}
library(macleish)
ggplot(data = whately_2015, aes(x = when, y = temperature)) + 
  geom_line(color = "darkgray") + 
  geom_smooth() + 
  xlab(NULL) + 
  ylab("Temperature (degrees Celsius)")
```
```{r}
# whately_2015 %>%
#   mutate(month = as.factor(lubridate::month(when, label = TRUE))) %>%
#   group_by(month) %>% 
#   skim(temperature) %>%
#   select(-na)

whately_2015$month <-
  as.factor(lubridate::month(whately_2015$when, label = TRUE))


#Base R
boxplot(
  whately_2015$temperature ~ whately_2015$month,
  xlab = "Month",
  ylab = expression("Temperature (" *  ~ degree * C * ")")
)

#ggplot
ggplot(
  data = whately_2015, 
  aes(
    x = lubridate::month(when, label = TRUE), 
    y = temperature
  )
) + 
  geom_boxplot() +
  xlab("Month") + 
  ylab(expression("Temperature ("*~degree*C*")"))
```


Check out the extended example here: https://mdsr-book.github.io/mdsr2e/ch-vizII.html#sec:babynames