---
title: "Iteration"
author: "Gregory J. Matthews"
date: "10/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Vectoring Operations
```{r}
#Let's say I want the averages of a bunch of columns. 
# I COULD use a for loop.
library(Lahman)
averages <- NULL
for (i in 15:40) {
  averages[i - 14] <- mean(Teams[, i], na.rm = TRUE)
}
names(averages) <- names(Teams)[15:40]
averages

```
"It is important to understand that the fundamental architecture of R is based on vectors. That is, in contrast to general-purpose programming languages like C++ or Python that distinguish between single items—like strings and integers—and arrays of those items, in R a “string” is just a character vector of length 1. There is no special kind of atomic object. Thus, if you assign a single “string” to an object, R still stores it as a vector." - Baumer Book
```{r}
a <- "a string"
class(a)

is.vector(a)

length(a)

```

Advice: "Try to avoid writing for() loops, even when it seems like the easiest solution."
