---
title: "regex"
author: "Gregory J. Matthews"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Strings
## Making strings
```{r}
string1 <- "This is a string"
string2 <- 'You can do it with single "quotes" too' 

string1
string2

#Note:
writeLines(string2)
```


```{r}
x <- c("\"", "\\")
x
#What's really in there
writeLines(x)
```


```{r}
#special characters
#tab
"\t"
#newline
"\n"

#Example Haiku
aging <- "Aging Curve, please end\nSomeday it will be published\nMy god end this  please"

aging
writeLines(aging)
```
## Vectors of strings
```{r}
#vector of strings
numbers <- c("one","two","three")

library(stringr)
#base R
nchar(numbers)
#stringr version
str_length(numbers)

#Base R paste
paste0("x","y")

#Combining strings
str_c("x", "y")

#Recycling
paste0("prefix-", c("a", "b", "c"), "-suffix")
str_c("prefix-", c("a", "b", "c"), "-suffix")

#Tan Ho example 
#This recycles
paste0("x",1:5,"y",1:3)
#this doesn't recycle
#str_c("x",1:5,"y",1:3)


##Collapse
#Base R
paste0(c("x","y"), collapse = "")
#tidyverse
str_c(c("x", "y"), collapse = "")

```

## Subsetting strings

```{r}
x <- c("Apple", "Banana", "Pear")
substring(x, 1, 3)
#Substring doesn't work with negative indices (indexes?)
substring(x, -3, -1)
#But this does!!!
substring(x, nchar(x) - 2, nchar(x))

#Tidyverse
str_sub(x, 1, 3)
#negative numbers count backwards from end
str_sub(x, -3, -1)

#what happens here?
substring("a",1,5)
str_sub("a",1,5)

```
# Regular Expressions
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view_all(x, "an")

#using . to match any character
str_view(x, ".a.")

#How do I find the actual .
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)

# And this tells R to look for an explicit .
str_view(c("abc", "a.c", "bef"), "a.c")
str_view(c("abc", "a.c", "bef"), "a\\.c")


#What does this match?  
#"\..\..\.."

```

## Anchors 

```{r}
#Match at the beginning with ^
#Match at the end with $
x <- c("apple", "banana", "pear")
str_view(x, "^a")
str_view(x, "a$")

#To remember which is which, try this mnemonic which I learned from Evan Misshula: if you begin with power (^), you end up with money ($).


#Anchor at beginning and end
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
str_view(x, "^apple$")

#Now you try: 
#Get all the words that being with y
#All the words that end with x
#All the words that are exactly 3 letters long (don't use str_length!)
#All the words with 7 or more letters
str_view(stringr::words, "regex_here")


```

## Character classes and alternatives
 - \\d: matches any digit. Also \[0-9] or \[:digits:]
 - \\s: matches any whitespace (e.g. space, tab, newline).
 - \[abc]: matches a, b, or c.
 - \[^abc]: matches anything except a, b, or c.

```{r}
#Instead of backslach you have this option
## Look for a literal character that normally has special meaning in a regex
str_view(c("abc", "a.c", "a*c", "a c"), "a[.]c")
str_view(c("abc", "a.c", "a*c", "a c"), ".[*]c")
str_view(c("abc", "a.c", "a*c", "a c"), "a[ ]")

#These must ALWAYS bw handled with back slash escpaes: ] \ ^ -

#Alternation
str_view(c("grey", "gray"), "gr(e|a)y")


#Now you try"
#Start with a vowel.
str_view(words, "^[aeiou]")
#That only contain consonants. (Hint: thinking about matching "not"-vowels.)
str_view(words, "[aeiou]", match = FALSE)
#End with "ed", but not with "eed".
str_view(words, "[abcdfghijklmnopqrstuvwxyz]ed$")
#End with "ing" or "ise".
str_view(words, "ing$|ise$")
#Empirically verify the rule “i before e except after c”.
str_view(words, "cie")
#Is "q" always followed by a "u"?
str_view(words, "q[letters[-21]]")
```


## Repetition
 - ?: 0 or 1
 - +: 1 or more
 - *: 0 or more
```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, 'C[LX]+')
str_view(x, 'C[LX]*')
```

You can also specify matches exactly: 
 - {n}: exactly n
 - {n,}: n or more
 - {,m}: at most m
 - {n,m}: between n and m
 
```{r}
str_view(x, "C{2}")
str_view(x, "C{2,}")
str_view(x, "C{2,3}")
```

Just a note: "By default these matches are “greedy”: they will match the longest string possible. You can make them “lazy”, matching the shortest string possible by putting a ? after them. This is an advanced feature of regular expressions, but it’s useful to know that it exists."  - https://r4ds.had.co.nz/strings.html


```{r}
#Now you try
#What will these match? 
#^.*$
#\d{4}-\d{2}-\d{2}

#Exercises
#Create regular expressions to find all words that:
#1. Start with three consonants.
#2. Have three or more vowels in a row.
library(babynames) 
library(tidyverse)
nam <- babynames %>% distinct(name)  %>% pull(name) %>% str_to_lower()

#1. Find all the names that end with xx or zz
#2. Find all the names that contain either the word deer or sleigh
#3. Find all the names containing rudolf or rudolph
#4. Find all the names containing santa or claus
#5. Find all the names that contain "sant" followed immediately by a vowel and the name ends in a vowel and begins with a consonant.  



#Note: To return a logical vector: str_detect


```




## Grouping and backreferences

"Parentheses also create a numbered capturing group (number 1, 2 etc.). A capturing group stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with backreferences, like \1, \2 etc."  - Hadley Wickham
 
```{r}
library(stringr)
str_view(fruit, "(..)\\1", match = TRUE)

#Three in a row
str_view(nam,"(.)\\1\\1")

#What will this do?  
str_view(nam,"(.).\\1.\\1")

#what will this do?  
str_view(nam,"(.)(.)(.).*\\3\\2\\1")
#How about this?  
str_view(nam,"(.)(.)(.).*\\1\\2\\3")

#Exercises
#Find names that start and end with the same letter. 
#Names that contain a repeated pair or letters. 
#Contains one letter repeated in at least three places.  
```

## Tools 


```{r}
#str_detect: returns a logical vector
#str_count: Counts the number of matches
mean(str_detect(words, "[aeiou]$"))
mean(str_detect(nam, "[aeiou]$"))
nam[str_detect(nam,"greg")]

x <- c("apple", "banana", "pear")
str_count(x, "a")

#Note: matches never overlap: 
str_count("abababa", "aba")
str_count_all("abababa", "aba")


#Lets name up some more questions with the names data set!
```





