---
title: "Tidy Data"
author: "Gregory J. Matthews"
date: "9/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidy Data

Data are called ``tidy" if they are organized according to two simple rules: 
 - The rows, called cases or observations, each refer to a specific, unique, and similar sort of thing, e.g., girls named Somaly in 1984.
 - The columns, called variables, each have the same sort of value recorded for each row. For instance, n gives the number of babies for each case; sex tells which gender was assigned at birth.
 
## Reshaping Data
```{r}
library(mdsr)
library(tidyverse)
BP_wide <- data.frame(subject = c("BHO","GWB","WJC"),
                      before = c(160,120, 105),
                      after = c(115, 135, 145))
#in wide format
BP_wide
```

```{r}
#In long format
BP_narrow <- data.frame(subject =  c("BHO","GWB","WJC","BHO","GWB","WJC"),
                      when = c(rep("before",3),rep("after",3)),
                     sbp = c(160,120,105,115,135,145))
BP_narrow                
```


#Pivoting wider
- The values_from argument is the name of the variable in the narrow format that is to be divided up into multiple variables in the resulting wide format. 

- The names_from argument is the name of the variable in the narrow format that identifies for each case individually which column in the wide format will receive the value.
```{r}
BP_narrow %>% 
  pivot_wider(names_from = when, values_from = sbp)
```

## Pivoting longer
```{r}
BP_wide %>% 
  pivot_longer(-subject, names_to = "when", values_to = "sbp")

BP_wide %>% 
  pivot_longer(c(before, after), names_to = "when", values_to = "sbp")
```


## Example 
In class exercises: 
 - Find the most gender neutral name
 - What name saw the biggest drop from 1900 to 1950?  How about from 1950 to 2000?  
```{r}

library(babynames)
babynames
#write.csv(babynames, file = "./data/babynames.csv", row.names = FALSE)
```


```{r}
library(tidyverse)
test <- babynames %>% group_by(name,sex) %>% summarise(total = sum(n)) %>% pivot_wider(names_from = sex, values_from = total) %>% mutate(ratio = M/(M+F)) %>% mutate(balanced = abs(ratio - .5)) %>% filter(M > 1000 & F > 1000) %>% arrange(balanced) 


```




